Faithful Variable Screening for\\ High-Dimensional Convex Regression

### 2015-05-02 tex/scam/reviews-v1/response-v1.tex

\item \rc{I really did not understand what was represented in figures 5(a) and
5(b) (axis?). Besides it does not seem to correspond with what is
described at the bottom of page 26 and at the top of page 27. The
notation $\|\cdot\|_{\infty,1}$ is not defined.}

We have further explained the old figure 5a and 5b (now figure 7a and 7b) in the experiment section. The following has been added to the paper:

... we plot on the y-axis the norm $\|f^{(t)}_j\|_{\infty}$ of every column $j$ against the regularization strength $\lambda^{(t)}$. Instead of plotting the value of $\lambda^{(t)}$ on the x-axis however, we plot the total norm at $\lambda^{(t)}$ normalized against the total norm at $\lambda^{(1)}$: $\frac{\sum_j \|f^{(t)}_j\|_{\infty}}{\sum_j \|f^{(1)}\|_{\infty}}$. Thus, as x-axis moves from 0 to 1, the regularization goes from strong to weak.

\item \rc{On page 31, at the top of the page, ``see discussion at beginning of
Section 5'' : I think it should be ``at the beginning''. In (8.1),
same problem as before with the indices : ``$\nu_{ki} d_{ki}$ '' must be replaced
with ``$\nu_{ik}d_{ik}$'' (see for instance page 20). At the bottom of the page,
$v_{ki}$ must be replaced by $\nu_{ki}$.}

The typo and the index notation inconsistency have been corrected.

@@

\item \rc{On page 32, on the first line of the second paragraph, ``$d_k = 0 \mbox{for $k
\in S$}$'', $S$ must be replaced by $S^c$ and $\|\mu\|$ by $\|\mu\|_1$ and ``It clear
that''. Next, the third paragraph (``to ease notational ....'') should
be placed at the beginning of the proof (it is used from the beginning
of the proof, cf conditions on $d_k$).}

We cannot move the third paragraph to the beginning because it is not clear at that point that $k$ is fixed. We have corrected all the notations before the third paragraph so that the notations are consistent.

\item \rc{On page 33, I find that $[\lambda\Delta_k^Tu]_1 = \lambda((X_{kn} −X_{k1})\kappa$. ( If $\kappa$ is be defined
in the following way : $\kappa = \frac{1}{\lambda_n(X_{kn}-X_{k1})}[\Delta_k^T u]_1$, then everything
goes well).}

The original definition of $\kappa$ was indeed wrong; we forgot to update $\kappa$ when we edited the proof. We have corrected this in the paper.

@@

We have corrected the instances in which we erroneously omitted ``... components bounded by $B$''.

\item \rc{On page 49, ``$\int p(x)^2dx \leq (\int p(x)dx)^2$ ``??}

We have corrected this mistake by placing an upper bound $c_u$ on the density $p(\mathbf{x})$ and incorporating that upper bound into this derivation. We now use the correct bound $\sqrt{ \int p(x)^2 dx } \leq c_u$.

\item \rc{On page 50, a constant 2 is missing in $\epsilon_n$.}

This mistake has been corrected across the proofs.

\item \rc{On page 41, in the second paragraph, I think the bound on $\sup_{h_L} |⟨w,h_L⟩_n|$ is
wrong. I would have used the fact that the variables $h_L(X_i) W_i$ are
independent, centered and sub-Gaussian with a scale smaller than
$2\sigma s B$. It gives a bound of order $sB\sigma \sqrt{\frac{\log\frac{2}{\delta}}{n}}$.}

@@

\subsection*{Reviewer 3}

Major comments:

\begin{enumerate}
\item \rc{The basis of all the results seem to be the boundary flatness
condition. This condition seems to be an artifact of the analysis and
based on my reading, it is unclear how restrctive this assumption
is. The authors claim it is weak just after stating it and then
present several straightforward examples in which it doesn't hold
(e.g. $f(x_1,x_2)=x_1 x_2$) and multivariate Gaussian examples. Hence it is unclear to me how
likely a given function is likely to fit into this framework.}

The boundary flatness condition is not restrictive in the sense that for any density $p(x)$ supported on $[0,1]^p$, and for any $\epsilon > 0$, there exists a boundary flat density whose $L_1$ distance with $p(x)$ is at most $\epsilon$ (Please see the newly added Example 3.2). This is because boundary flatness places no restriction on what $p(x)$ looks like in the interior of the hypercube. Boundary flatness allows arbitrary correlation structure, provided that $p(x) > 0$. We have added Section 3.2 to the paper to give a more detailed discussion on boundary flatness.

$f(x_1,x_2) = x_1 x_2$ is non-convex and it is an example of how non-convex functions may not be additively faithful under the uniform distribution; it is not an example in which boundary flatness doesn't hold. The multivariate Gaussian is a counterexample but it is to show that boundary flatness may not extend easily to densities with unbounded support.

We acknowledge that many distributions are not boundary flat. We present boundary flatness as a significant generalization of the product density, under which nonparametric variable selection is already very difficult.

Boundary flatness is not a contrived condition. We initially proved additive faithfulness only for product densities and then went back and gradually weakened the condition on the density over many iterations of refinement.

### 2015-05-02 tex/scam/reviews-v1/response-v1.tex

\begin{enumerate}
\item \rc{Boundary flatness condition: much of the paper's analysis hinges on
this condition.  Reviewer 3 raises some concerns about when it holds,
especially for functions that are not necessarily supported on the
unit hypercube.  See also the point below.}

We have added commentary on this here and here.

\item \rc{Intuitively, variable selection is very closely related to
estimating a function (or possibly its partial derivatives) in the
sup-norm.  It is well-known that accurate estimation of functions
(whether in $L^2$ or the sup-norm) with bounded smoothness in dimension
$s$ requires a sample size that grows exponentially in $s$.  For this
reason, the past result of Comminges and Dalalyan (suggesting an
exponential growth for variable selection) is to be expected.
However, in terms of metric entropy, the class of convex functions
behaves roughly like the class of two-smooth functions (e.g., Dryanov
(2009)).  So at first glance, this makes the result of this paper
rather surprising---namely, that one can somehow side-step any
exponential growth in $s$, even though the metric entropy is still
scaling as badly as the case of 2-smooth functions.  I wondered if the
authors could comment and shed insight on this fact, which seems a
little counterintuitive.  What I suspect that it could mean is that
the additive faithfulness (AF) condition, or one of the sufficient
conditions for AF (namely boundary flatness) are somehow very
restrictive.  Could the authors comment on these issues?}

The boundary flatness doesn't artificially make the variable selection easy. Boundary flatness is much more general than the product density and Comminges and Dalalyan showed that, even under a product density, variable selection for general two-smooth functions is hard. Any bounded density with compact support can be approximated arbitrary well by a boundary flat density (see the newly added Example 3.2).

The intuition behind our result is the following: start with the fact that convex functions are additively faithful under a product density (see paragraphs below Theorem 3.1 for an intuition of this). The behavior of a convex function everywhere is constrained by its behavior at the boundary. Therefore, it is sufficient that the underlying density resembles a product density only at the boundary, which is exactly the notion that boundary flatness formalizes.

What we think is that, for functions with shape-constraints such as convexity, the metric entropy alone does not determine the hardness of variable selection; the shape-constraint too plays a direct role.

### 2015-05-02 tex/scam/reviews-v1/response-v1.tex

The boundary flatness condition is not restrictive in the sense that for any density $p(x)$ supported on $[0,1]^p$, and for any $\epsilon > 0$, there exists a boundary flat density whose $L_1$ distance with $p(x)$ is at most $\epsilon$ (Please see the newly added Example 3.2). This is because boundary flatness places no restriction on what $p(x)$ looks like in the interior of the hypercube. Boundary flatness allows arbitrary correlation structure, provided that $p(x) > 0$. We have added Section 3.2 to the paper to give a more detailed discussion on boundary flatness.

$f(x_1,x_2) = x_1 x_2$ is non-convex and it is an example of how non-convex functions may not be additively faithful under the uniform distribution; it is not an example in which boundary flatness doesn't hold. The multivariate Gaussian is a counterexample but it is to show that boundary flatness may not extend easily to densities with unbounded support.

We acknowledge that many distributions are not boundary flat. We present boundary flatness as a significant generalization of the product density, under which nonparametric variable selection is already very difficult.

Boundary flatness is not a contrived condition. We initially proved additive faithfulness only for product densities. But then, knowing that the behavior of a convex function everywhere is constrained by its behavior at the boundary, we conjectured that underlying density may only need to ``look like'' a product density at the boundary. Attempts to formalize this notion led us to the definition of boundary flatness.

### 2015-05-02 tex/scam/reviews-v1/response-v1.tex

We have added commentary on this here and here.

\item \rc{Intuitively, variable selection is very closely related to
estimating a function (or possibly its partial derivatives) in the
sup-norm.  It is well-known that accurate estimation of functions
(whether in $L^2$ or the sup-norm) with bounded smoothness in dimension
$s$ requires a sample size that grows exponentially in $s$.  For this
reason, the past result of Comminges and Dalalyan (suggesting an
exponential growth for variable selection) is to be expected.
However, in terms of metric entropy, the class of convex functions
behaves roughly like the class of two-smooth functions (e.g., Dryanov
(2009)).  So at first glance, this makes the result of this paper
rather surprising---namely, that one can somehow side-step any
exponential growth in $s$, even though the metric entropy is still
scaling as badly as the case of 2-smooth functions.  I wondered if the
authors could comment and shed insight on this fact, which seems a
little counterintuitive.  What I suspect that it could mean is that
the additive faithfulness (AF) condition, or one of the sufficient
conditions for AF (namely boundary flatness) are somehow very
restrictive.  Could the authors comment on these issues?}

The boundary flatness doesn't artificially make the variable selection easy. Boundary flatness is much more general than the product density and Comminges and Dalalyan showed that, even under a product density, variable selection for general two-smooth functions is hard. Any bounded density with compact support can be approximated arbitrary well by a boundary flat density (see the newly added Example 3.2).

The intuition behind our result is the following: start with the fact that convex functions are additively faithful under a product density (see paragraphs below Theorem 3.1 for an intuition of this). The behavior of a convex function everywhere is constrained by its behavior at the boundary. Therefore, we need only that the underlying density resembles a product density at the boundary, which is exactly the notion that boundary flatness formalizes.

### 2015-05-02 tex/scam/reviews-v1/response-v1.tex

Smoothness can be incorporated by using B-spline basis with uniformly spaced knots. The details are described in a 2014 paper by Pya and Wood which we have cited in the introduction. We chose not to discuss estimating smooth convex functions because it would introduce an additional tuning parameter.

The analysis may require much more work if one uses a B-spline basis in the estimation. B-splines may significantly complicate the false positive analysis, which rely on the particular form of the optimization. The false negative analysis may not be affected much because it mainly uses bracketing entropy.

\item \rc{Assuming model selection consistency, can one derive prediction
error bounds using the fitted $f_k$'s and $g_k$'s, as well? Prediction error
bounds seem quite relevant due to possible model misspecification when
$f_0$ is not actually a sum of univariate convex functions.}

To minimize predictive error, we can refit a non-additive convex function after selecting a sparse model. Comparing the predictive accuracy of $f^*_k$'s and $g^*_k$'s against the non-additive true regression function $f_0$ is beyond the scope of our paper, but we do compare the finite sample estimate $\hat{f}_k$ and $\hat{g}_k$ against the population level $f^*_k$ and $g^*_k$. The analysis we do to control false negatives also shows that $$\left| \mathbb{E}\left( f_0(X)  - \sum_{k=1}^p f^*_k(X_k) \right)^2 - \mathbb{E} \left( f_0(X) - \sum_{k=1}^p \hat{f}_k(X_k))\right)^2 \right| \rightarrow 0$$ and likewise for the $\hat{g}_k$'s. Please see Theorem 8.3 and 8.4 in the supplement.

@@

\item \rc{In order to improve clarity, the authors should define the convex
additive model at the top of page 5: $f(x_i) = \sum_{k=1}^p f_k(x_{ik})$, where the
$f_k$'s are convex.  Otherwise, the notation in equation (2.3) is
initially a bit confusing.}

We have removed the old equation (2.3) and any discussion of convex additive model at the top of page 5. Everything there is discussed on page 6, after AC/DC has been introduced.

@@

\item \rc{The term $\epsilon$ supposed to balance the two terms is not correct (for
instance it does not contain $\delta$). There are many small mistakes in the
calculations. In particular, the exponent in $s$ is not the right one.}

The $\epsilon$ we choose is sufficient to get the bounds we want but it does not exactly balance the terms. We left dependence on $\delta$ out of the $\epsilon$ intentionally because we need to verify that $\epsilon$ satisfies certain conditions, see Equation (8.12) for instance. These conditions are much easier to verify if $\epsilon$ does not depend on $\delta$, which we set to be $1/n$ only at the end. We have updated the paper to clearly state that we use a suboptimal $\epsilon$.

\item \rc{In the proof of Theorem 8.4 : too many small mistakes in
calculations, but the final result is OK.}

We have carefully proof-read and edited the proof of Theorem 8.4.

@@

\end{enumerate}

Minor comments:

\begin{enumerate}
\item \rc{I didn't think there was any need for Section 2
  since it feels very repetitive and slightly confusing that parts of
  the text are repeated almost verbatim. I would suggest removing this section.}

Section 2 is repetitive but we believe it has value in highlighting the big picture, which could otherwise be overwhelmed by the technical details.

\item \rc{Page 2, second last paragraph `and and.'}

We have corrected this.

### 2015-05-02 tex/scam/estimation.tex
minor edit
major paper revisions
more response

Theorem~\ref{thm:acdc_faithful} naturally suggests 
a two-stage screening procedure for variable selection in the population setting. In the first stage, we fit a convex additive model. 
\begin{equation}
\label{eqn:scam2_pop}
f^*_1, ..., f^*_p = \argmin_{f_1,...,f_p \in \C^1_0} 
   \E \Big( f(X)  - \sum_{k=1}^p f_k(X_k) \Big)^2 
\end{equation}
where we denote $\C^1_0$ ($\mh{}\C^1_0$) as the set of one-dimensional convex (resp. concave) functions with population mean zero. In the second stage, for every variable marked as irrelevant in the first stage, we fit a univariate \emph{concave} function separately on the residual for that variable.
 for each $k$ such that $ f^*_k = 0$:
\begin{equation}
\label{eqn:dc2}
g^*_k = \argmin_{g_k \in \mh{}\C^1_0} 
   \E \Big( f(X) - \sum_{k'} f^*_{k'}(X_{k'}) 
    - g_k(X_{k})\Big)^2 
\end{equation}
We screen out $S^C$, any variable $k$ that is zero after the second stage, and output $S$.
\begin{equation}
\label{eqn:acdc_vars_pop}
S^c = \bigl\{k : f^*_k =
0 \; \mathrm{and}\; g^*_k =0\bigr\}.
\end{equation}

@@

@@

\begin{figure}[t]
{\sc AC/DC Algorithm for Variable Selection in Convex Regression\hfill}
\vskip5pt
\begin{center}
\hrule
\vskip7pt
\normalsize
\begin{enumerate}
\item[] \textit{Input}:  $(\mathbf{x}_1, y_1), ..., (\mathbf{x}_n, y_n)$, regularization parameter $\lambda$.
\vskip5pt
\item[] \textit{AC Stage}:  Estimate a sparse additive convex model:
\begin{equation}
\label{eqn:scam2}
\hat{f}_1, ..., \hat{f}_p, \hat\mu = \argmin_{f_1,...,f_p \in \C^1_0} 
   \frac{1}{n} \sum_{i=1}^n \Big(y_i - \mu-\sum_{k=1}^p f_k(x_{ik}) \Big)^2 
       + \lambda \sum_{k=1}^p \| f_k \|_\infty
\end{equation}
\vskip5pt
\item[] \textit{DC Stage}:  Estimate concave functions
 for each $k$ such that $\| \hat{f}_k \|_\infty = 0$:
\begin{equation}
\label{eqn:dc2}
\hat{g}_k = \argmin_{g_k \in \mh{}\C^1_0} 
   \frac{1}{n} \sum_{i=1}^n \Big( y_i - \hat \mu - \sum_{k'} \hat{f}_{k'}(x_{ik'}) 
    - g_k(x_{ik})\Big)^2 
      + \lambda \| g_k \|_\infty
\end{equation}
\item[] \textit{Output}: Component functions $\{\hat f_k\}$ and 
relevant variables $\hat S$ where
\begin{equation}
\hat S^c = \bigl\{k : \| \hat{f}_k \| =
0 \; \mathrm{and}\; \|\hat{g}_k \|=0\bigr\}.
\end{equation}
\end{enumerate}
\vskip3pt
\hrule
\end{center}
\vskip0pt
\caption{The AC/DC algorithm for variable selection in convex
  regression.  The AC stage fits a sparse additive convex regression
  model, using a quadratic program that imposes an group sparsity
  penalty for each component function.  The DC stage fits
  decoupled concave functions on the residuals, for each 
  component that is zeroed out in the AC stage.}
\label{fig:backfitting:algo}
\end{figure}

>>>>

Theorem~\ref{thm:acdc_faithful} naturally suggests 
a two-stage screening procedure for variable selection in the population setting. In the first stage, we fit a convex additive model. 
\begin{equation}
\label{eqn:scam2_pop}
f^*_1, ..., f^*_p = \argmin_{f_1,...,f_p \in \C^1_0, \mu} 
   \E \Big( f(X) - \mu - \sum_{k=1}^p f_k(X_k) \Big)^2 
\end{equation}
where we denote $\C^1_0$ ($\mh{}\C^1_0$) as the set of one-dimensional convex (resp. concave) functions with population mean zero. In the second stage, for every variable marked as irrelevant in the first stage, we fit a univariate \emph{concave} function separately on the residual for that variable.
 For each $k$ such that $ f^*_k = 0$:
\begin{equation}
\label{eqn:dc2}
g^*_k = \argmin_{g_k \in \mh{}\C^1_0} 
   \E \Big( f(X) - \mu^* - \sum_{k'} f^*_{k'}(X_{k'}) 
    - g_k(X_{k})\Big)^2 
\end{equation}
We screen out $S^C$, any variable $k$ that is zero after the second stage, and output $S$.
\begin{equation}
\label{eqn:acdc_vars_pop}
S^c = \bigl\{k : f^*_k =
0 \; \mathrm{and}\; g^*_k =0\bigr\}.
\end{equation}

@@

After selecting the variable set $\hat{S}$, one can refit a low-dimensional non-additive convex function to build the best predictive model. If refitting is undesirable for whatever reason, the AC/DC outputs can also be used for prediction. Given a new sample $\mathbf{x}$, we let $\hat{y} = \sum_k \hat{f}_k(\mathbf{x}_k) + \sum_k \hat{g}_k(\mathbf{x}_k)$. Note that $\hat{g}_k = 0$ for $k$ such that $\hat{f}_k \neq 0$ in AC/DC. The next section describes how to compute this function evaluation.

@@

\begin{figure}[t]
{\sc AC/DC Algorithm for Variable Selection in Convex Regression\hfill}
\vskip5pt
\begin{center}
\hrule
\vskip7pt
\normalsize
\begin{enumerate}
\item[] \textit{Input}:  $(\mathbf{x}_1, y_1), ..., (\mathbf{x}_n, y_n)$, regularization parameter $\lambda$.
\vskip5pt
\item[] \textit{AC Stage}:  Estimate a sparse additive convex model:
\begin{equation}
\label{eqn:scam2}
\hat{f}_1, ..., \hat{f}_p, \hat\mu = \argmin_{f_1,...,f_p \in \C^1_0} 
   \frac{1}{n} \sum_{i=1}^n \Big(y_i - \mu-\sum_{k=1}^p f_k(x_{ik}) \Big)^2 
       + \lambda \sum_{k=1}^p \| f_k \|_\infty
\end{equation}
\vskip5pt
\item[] \textit{DC Stage}:  Estimate concave functions
 for each $k$ such that $\| \hat{f}_k \|_\infty = 0$:
\begin{equation}
\label{eqn:dc22}
\hat{g}_k = \argmin_{g_k \in \mh{}\C^1_0} 
   \frac{1}{n} \sum_{i=1}^n \Big( y_i - \hat \mu - \sum_{k'} \hat{f}_{k'}(x_{ik'}) 
    - g_k(x_{ik})\Big)^2 
      + \lambda \| g_k \|_\infty
\end{equation}
\item[] \textit{Output}: Component functions $\{\hat f_k\}$ and 
relevant variables $\hat S$ where
\begin{equation}
\hat S^c = \bigl\{k : \| \hat{f}_k \| =
0 \; \mathrm{and}\; \|\hat{g}_k \|=0\bigr\}.
\end{equation}
\end{enumerate}
\vskip3pt
\hrule
\end{center}
\vskip0pt
\caption{The AC/DC algorithm for variable selection in convex
  regression.  The AC stage fits a sparse additive convex regression
  model, using a quadratic program that imposes an group sparsity
  penalty for each component function.  The DC stage fits
  decoupled concave functions on the residuals, for each 
  component that is zeroed out in the AC stage.}
\label{fig:backfitting:algo}
\end{figure}
